---
title: "main"
author: "Jing Zhao (jz2786)"
date: "11/28/2017"
output: pdf_document
---

# Memory-Based Algorithm

## Load Data
```{r}
movie_train <- read.csv("../data/data_sample/eachmovie_sample/data_train.csv")
movie_test <- read.csv("../data/data_sample/eachmovie_sample/data_test.csv")
MS_train <- read.csv("../data/data_sample/MS_sample/data_train.csv")
MS_test <- read.csv("../data/data_sample/MS_sample/data_test.csv")
```

## Transformation
Convert the original dataset to a matrix which rows represents users and columns represents items
For dataset 1 (MS), we assign 0 to those items which users never clicked.
For dataset 2 (Movie), we assign NA to those items which users never rated.
```{r}
source("../lib/memoryBased.R")
movie_train <- Transformer(movie_train)
movie_test <- Transformer(movie_test)
#save(movie_train, file = "../output/movie_train.RData")
#save(movie_test, file = "../output/movie_test.RData")

MS_train <- Transformer2(MS_train)
MS_test <- Transformer2(MS_test)
#save(MS_train, file = "../output/MS_train.RData")
#save(MS_test, file = "../output/MS_test.RData")
```

## Calculate similarity weights 

Pearson Correlation & Spearman Correlation & Vector Similarity
```{r}

# Dataset 2 (Movie)

# pearson_weight <- cal_weight(movie_train,'pearson')
# spearman_weight <- cal_weight(movie_train,'spearman')
# vector_weight <- cal_weight(movie_train,'vector')
# 
# pearson_weight[is.na(pearson_weight)] = 0
# spearman_weight[is.na(spearman_weight)] = 0
# vector_weight[is.na(vector_weight)] = 0
#  
# save(pearson_weight,file = './pearson_weight.RData')
# save(spearman_weight,file='./spearman_weight.RData')
# save(vector_weight,file = './vector_weight.RData')
 
# Dataset 1 (MS)

# pearson_weight_MS <- cal_weight(MS_train,'pearson')
# spearman_weight_MS <- cal_weight(MS_train,'spearman')
# vector_weight_MS <- cal_weight(MS_train,'vector')
# 
# pearson_weight_MS[is.na(pearson_weight_MS)] = 0
# spearman_weight_MS[is.na(spearman_weight_MS)] = 0
# vector_weight_MS[is.na(vector_weight_MS)] = 0
#  
# save(pearson_weight_MS,file = './pearson_weight_MS.RData')
# save(spearman_weight_MS,file='./spearman_weight_MS.RData')
# save(vector_weight_MS,file = './vector_weight_MS.RData')


```

## Significance Weighting
```{r}
# Dataset 2 (Movie)

# mat_significance_weight_movie <- significance_weight_matrix(dim(movie_train)[1], mat = movie_train)
# mat_significance_weight_movie[is.na(mat_significance_weight_movie)] = 0
# save(mat_significance_weight_movie, file = "significance_weight.RData")

# Dataset 1 (MS)

# mat_significance_weight_MS <- significance_weight_matrix(dim(MS_train)[1], mat = MS_train)
# mat_significance_weight_MS[is.na(mat_significance_weight_MS)] = 0
# save(mat_significance_weight_MS, file = "significance_weight2.RData")
```

## Variance Weighting
```{r}
# Dataset 2 (Movie)

# mat_variance_weight_movie <- variance_weight_matrix(dim(movie_train)[1], mat = movie_train)
# mat_variance_weight_movie[is.na(mat_variance_weight_movie)] = 0
# save(mat_variance_weight_movie, file = "variance_weight.RData")

# Dataset 1 (MS)

# mat_variance_weight_MS <- variance_weight_matrix(dim(MS_train)[1], mat = MS_train)
# mat_variance_weight_MS[is.na(mat_variance_weight_MS)] = 0
# save(mat_variance_weight_MS, file = "variance_weight2.RData")
```

```{r}
load('../output/similarity_weight/pearson_weight.RData')
load('../output/similarity_weight/pearson_weight_MS.RData')
load('../output/similarity_weight/spearman_weight.RData')
load('../output/similarity_weight/spearman_weight_MS.RData')
load('../output/similarity_weight/vector_weight.RData')
load('../output/similarity_weight/vector_weight_MS.RData')
load('../output/similarity_weight/simrank_weight.RData')
simrank_weight[is.na(simrank_weight)] <- 0
mat_significance_weight_movie <- get(load('../output/significance_weight/significance_weight.RData'))
mat_significance_weight_MS <-get(load('../output/significance_weight/significance_weight2.RData'))
mat_variance_weight_movie <- get(load('../output/variance_weight/variance_weight.RData'))
mat_variance_weight_MS <- get(load('../output/variance_weight/variance_weight2.RData'))
mat_significance_weight_movie[is.na(mat_significance_weight_movie)] = 0
mat_significance_weight_MS[is.na(mat_significance_weight_MS)] = 0
mat_variance_weight_movie[is.na(mat_variance_weight_movie)] = 0
mat_variance_weight_MS[is.na(mat_variance_weight_MS)] = 0
```

## Neighbors Selection & Rating Normalization
Neighbors Selection Methods: Weight Threshold
Rating Normalization Methods: Deviation for Mean
```{r}
# Pearson
pearson.movie.neighbor = neighbors.select(pearson_weight)
pearson.MS.neighbor = neighbors.select(pearson_weight_MS)
pearson.movie.pred = pred.matrix.movie(simweights =pearson_weight, top.neighbors =pearson.movie.neighbor)
pearson.MS.pred = pred.matrix.ms(simweights =pearson_weight_MS, top.neighbors =pearson.MS.neighbor)

# Spearman 
spearman.movie.neighbor = neighbors.select(spearman_weight)
spearman.MS.neighbor = neighbors.select(spearman_weight_MS)
spearman.movie.pred = pred.matrix.movie(simweights =spearman_weight, top.neighbors =spearman.movie.neighbor)
spearman.MS.pred = pred.matrix.ms(simweights =spearman_weight_MS, top.neighbors =spearman.MS.neighbor)

# Vector 
vector.movie.neighbor = neighbors.select(vector_weight)
vector.MS.neighbor = neighbors.select(vector_weight_MS)
vector.movie.pred = pred.matrix.movie(simweights =vector_weight, top.neighbors =vector.movie.neighbor)
vector.MS.pred = pred.matrix.ms(simweights =vector_weight_MS, top.neighbors =vector.MS.neighbor)

# Pearson + Sig
pearson.sig.movie = mat_significance_weight_movie * pearson_weight
pearson.sig.MS = mat_significance_weight_MS * pearson_weight_MS
pearson.sig.movie.neighbor =  neighbors.select(pearson.sig.movie)
pearson.sig.MS.neighbor = neighbors.select(pearson.sig.MS, thresholding = 0.013)
pearson.sig.movie.pred = pred.matrix.movie(simweights = pearson.sig.movie, top.neighbors = pearson.sig.movie.neighbor)
pearson.sig.MS.pred = pred.matrix.ms(simweights = pearson.sig.MS, top.neighbors = pearson.sig.MS.neighbor)

# Spearman + Sig
spearman.sig.movie = mat_significance_weight_movie * spearman_weight
spearman.sig.MS = mat_significance_weight_MS * spearman_weight_MS
spearman.sig.movie.neighbor =  neighbors.select(spearman.sig.movie)
spearman.sig.MS.neighbor = neighbors.select(spearman.sig.MS, thresholding = 0.013)
spearman.sig.movie.pred = pred.matrix.movie(simweights =spearman.sig.movie,  top.neighbors = spearman.sig.movie.neighbor)
spearman.sig.MS.pred = pred.matrix.ms(simweights =spearman.sig.MS,  top.neighbors =spearman.sig.MS.neighbor)

# Vector + Sig
vector.sig.movie = mat_significance_weight_movie * vector_weight
vector.sig.MS = mat_significance_weight_MS * vector_weight_MS
vector.sig.movie.neighbor =  neighbors.select(vector.sig.movie)
vector.sig.MS.neighbor = neighbors.select(vector.sig.MS, thresholding = 0.014)
vector.sig.movie.pred = pred.matrix.movie(simweights =vector.sig.movie,top.neighbors = vector.sig.movie.neighbor)
vector.sig.MS.pred = pred.matrix.ms(simweights =vector.sig.MS, top.neighbors =vector.sig.MS.neighbor)

# Pearson + Var
pearson.var.movie = mat_variance_weight_movie * pearson_weight
pearson.var.MS = mat_variance_weight_MS * pearson_weight_MS
pearson.var.movie.neighbor =  neighbors.select(pearson.var.movie)
pearson.var.MS.neighbor = neighbors.select(pearson.var.MS)
pearson.var.movie.pred = pred.matrix.movie(simweights =pearson.var.movie,top.neighbors = pearson.var.movie.neighbor)
pearson.var.MS.pred = pred.matrix.ms(simweights =pearson.var.MS, top.neighbors =pearson.var.MS.neighbor)

# Spearman + Var
spearman.var.movie = mat_variance_weight_movie * spearman_weight
spearman.var.MS = mat_variance_weight_MS * spearman_weight_MS
spearman.var.movie.neighbor =  neighbors.select(spearman.var.movie)
spearman.var.MS.neighbor = neighbors.select(spearman.var.MS)
spearman.var.movie.pred = pred.matrix.movie(simweights =spearman.var.movie,top.neighbors = spearman.var.movie.neighbor)
spearman.var.MS.pred = pred.matrix.ms(simweights =spearman.var.MS,top.neighbors = spearman.var.MS.neighbor)

# Vector + Var
vector.var.movie = mat_variance_weight_movie * vector_weight
vector.var.MS = mat_variance_weight_MS * vector_weight_MS
vector.var.movie.neighbor =  neighbors.select(vector.var.movie)
vector.var.MS.neighbor = neighbors.select(vector.var.MS)
vector.var.movie.pred = pred.matrix.movie(simweights =vector.var.movie,top.neighbors = vector.var.movie.neighbor)
vector.var.MS.pred = pred.matrix.ms(simweights =vector.var.MS, top.neighbors =vector.var.MS.neighbor)

# Pearson + Sig + Var
pearson.sig.var.movie = pearson.sig.movie * mat_variance_weight_movie
pearson.sig.var.MS = pearson.sig.MS * mat_variance_weight_MS
pearson.sig.var.movie.neighbor =  neighbors.select(pearson.sig.var.movie)
pearson.sig.var.MS.neighbor = neighbors.select(pearson.sig.var.MS, thresholding = 0.048)
pearson.sig.var.movie.pred = pred.matrix.movie(simweights =pearson.sig.var.movie,top.neighbors = pearson.sig.var.movie.neighbor)
pearson.sig.var.MS.pred = pred.matrix.ms(simweights =pearson.sig.var.MS,top.neighbors = pearson.sig.var.MS.neighbor)

# Spearman + Sig + Var
spearman.sig.var.movie = spearman.sig.movie * mat_variance_weight_movie
spearman.sig.var.MS = spearman.sig.MS * mat_variance_weight_MS
spearman.sig.var.movie.neighbor =  neighbors.select(spearman.sig.var.movie)
spearman.sig.var.MS.neighbor = neighbors.select(spearman.sig.var.MS, thresholding = 0.048)
spearman.sig.var.movie.pred = pred.matrix.movie(simweights =spearman.sig.var.movie,top.neighbors = spearman.sig.var.movie.neighbor)
spearman.sig.var.MS.pred = pred.matrix.ms(simweights =spearman.sig.var.MS,top.neighbors = spearman.sig.var.MS.neighbor)

# Vector + Sig + Var
vector.sig.var.movie = vector.sig.movie * mat_variance_weight_movie
vector.sig.var.MS = vector.sig.MS * mat_variance_weight_MS
vector.sig.var.movie.neighbor =  neighbors.select(vector.sig.var.movie)
vector.sig.var.MS.neighbor = neighbors.select(vector.sig.var.MS, thresholding = 0.05)
vector.sig.var.movie.pred = pred.matrix.movie(simweights =vector.sig.var.movie,top.neighbors = vector.sig.var.movie.neighbor)
vector.sig.var.MS.pred = pred.matrix.ms(simweights =vector.sig.var.MS,top.neighbors = vector.sig.var.MS.neighbor)

# SimRank
simrank.movie.neighbor = neighbors.select(simrank_weight,thresholding = 0.005)
simrank.movie.pred = pred.matrix.movie(simweights =simrank_weight, top.neighbors =simrank.movie.neighbor)

# SimRank + Sig
simrank.sig.movie = mat_significance_weight_movie * simrank_weight
simrank.sig.movie.neighbor =  neighbors.select(simrank.sig.movie, thresholding = 0.005)
simrank.sig.movie.pred = pred.matrix.movie(simweights = simrank.sig.movie, top.neighbors = simrank.sig.movie.neighbor)

# SimRank + Var
simrank.var.movie = mat_variance_weight_movie * simrank_weight
simrank.var.movie.neighbor =  neighbors.select(simrank.var.movie, thresholding = 0.005)
simrank.var.movie.pred = pred.matrix.movie(simweights =simrank.var.movie,top.neighbors = simrank.var.movie.neighbor)

# SimRank + Sig + Var 
simrank.sig.var.movie = simrank.sig.movie * mat_variance_weight_movie
simrank.sig.var.movie.neighbor =  neighbors.select(simrank.sig.var.movie, thresholding = 0.005)
simrank.sig.var.movie.pred = pred.matrix.movie(simweights =simrank.sig.var.movie,top.neighbors = simrank.sig.var.movie.neighbor)

```

## Evaluation - MAE
```{r}
# Movie
pearson.movie.mae = evaluation.mae(pearson.movie.pred, movie_test)
pearson.sig.movie.mae = evaluation.mae(pearson.sig.movie.pred, movie_test)
pearson.var.movie.mae = evaluation.mae(pearson.var.movie.pred, movie_test)
pearson.sig.var.movie.mae = evaluation.mae(pearson.sig.var.movie.pred, movie_test)

spearman.movie.mae = evaluation.mae(spearman.movie.pred, movie_test)
spearman.sig.movie.mae = evaluation.mae(spearman.sig.movie.pred, movie_test)
spearman.var.movie.mae = evaluation.mae(spearman.var.movie.pred, movie_test)
spearman.sig.var.movie.mae = evaluation.mae(spearman.sig.var.movie.pred, movie_test)

vector.movie.mae = evaluation.mae(vector.movie.pred, movie_test)
vector.sig.movie.mae = evaluation.mae(vector.sig.movie.pred, movie_test)
vector.var.movie.mae = evaluation.mae(vector.var.movie.pred, movie_test)
vector.sig.var.movie.mae = evaluation.mae(vector.sig.var.movie.pred, movie_test)

simrank.movie.mae = evaluation.mae(simrank.movie.pred, movie_test)
simrank.sig.movie.mae = evaluation.mae(simrank.sig.movie.pred, movie_test)
simrank.var.movie.mae = evaluation.mae(simrank.var.movie.pred, movie_test)
simrank.sig.var.movie.mae = evaluation.mae(simrank.sig.var.movie.pred, movie_test)


pearson.movie.mae 
pearson.sig.movie.mae 
pearson.var.movie.mae 
pearson.sig.var.movie.mae 
spearman.movie.mae 
spearman.sig.movie.mae
spearman.var.movie.mae
spearman.sig.var.movie.mae 
vector.movie.mae 
vector.sig.movie.mae 
vector.var.movie.mae 
vector.sig.var.movie.mae 
simrank.movie.mae 
simrank.sig.movie.mae 
simrank.var.movie.mae 
simrank.sig.var.movie.mae 
```

## Evaluation - ROC
```{r}
# Movie
pearson.movie.roc = evaluation.roc(4, pearson.movie.pred, movie_test)
pearson.sig.movie.roc = evaluation.roc(4, pearson.sig.movie.pred, movie_test)
pearson.var.movie.roc = evaluation.roc(4, pearson.var.movie.pred, movie_test)
pearson.sig.var.movie.roc = evaluation.roc(4, pearson.sig.var.movie.pred, movie_test)

spearman.movie.roc = evaluation.roc(4, spearman.movie.pred, movie_test)
spearman.sig.movie.roc = evaluation.roc(4, spearman.sig.movie.pred, movie_test)
spearman.var.movie.roc = evaluation.roc(4, spearman.var.movie.pred, movie_test)
spearman.sig.var.movie.roc = evaluation.roc(4, spearman.sig.var.movie.pred, movie_test)

vector.movie.roc = evaluation.roc(4, vector.movie.pred, movie_test)
vector.sig.movie.roc = evaluation.roc(4, vector.sig.movie.pred, movie_test)
vector.var.movie.roc = evaluation.roc(4, vector.var.movie.pred, movie_test)
vector.sig.var.movie.roc = evaluation.roc(4, vector.sig.var.movie.pred, movie_test)

simrank.movie.roc = evaluation.roc(4, simrank.movie.pred, movie_test)
simrank.sig.movie.roc = evaluation.roc(4, simrank.sig.movie.pred, movie_test)
simrank.var.movie.roc = evaluation.roc(4, simrank.var.movie.pred, movie_test)
simrank.sig.var.movie.roc = evaluation.roc(4, simrank.sig.var.movie.pred, movie_test)

pearson.movie.roc 
pearson.sig.movie.roc 
pearson.var.movie.roc 
pearson.sig.var.movie.roc
spearman.movie.roc
spearman.sig.movie.roc 
spearman.var.movie.roc 
spearman.sig.var.movie.roc
vector.movie.roc 
vector.sig.movie.roc 
vector.var.movie.roc
vector.sig.var.movie.roc 
simrank.movie.roc 
simrank.sig.movie.roc 
simrank.var.movie.roc
simrank.sig.var.movie.roc 

```

## Evaluation - Rank Score
```{r}
# MS
pearson.MS.rankscore = rank_score(pearson.MS.pred, MS_test)
pearson.sig.MS.rankscore = rank_score(pearson.sig.MS.pred, MS_test)
pearson.var.MS.rankscore = rank_score(pearson.var.MS.pred, MS_test)
pearson.sig.var.MS.rankscore = rank_score(pearson.sig.var.MS.pred, MS_test)

spearman.MS.rankscore = rank_score(spearman.MS.pred, MS_test)
spearman.sig.MS.rankscore = rank_score(spearman.sig.MS.pred, MS_test)
spearman.var.MS.rankscore = rank_score(spearman.var.MS.pred, MS_test)
spearman.sig.var.MS.rankscore = rank_score(spearman.sig.var.MS.pred, MS_test)

vector.MS.rankscore = rank_score(vector.MS.pred, MS_test)
vector.sig.MS.rankscore = rank_score(vector.sig.MS.pred, MS_test)
vector.var.MS.rankscore = rank_score(vector.var.MS.pred, MS_test)
vector.sig.var.MS.rankscore = rank_score(vector.sig.var.MS.pred, MS_test)

pearson.MS.rankscore
pearson.sig.MS.rankscore 
pearson.var.MS.rankscore
pearson.sig.var.MS.rankscore
spearman.MS.rankscore 
spearman.sig.MS.rankscore 
spearman.var.MS.rankscore 
spearman.sig.var.MS.rankscore
vector.MS.rankscore
vector.sig.MS.rankscore 
vector.var.MS.rankscore 
vector.sig.var.MS.rankscore 

```

# Model-Based Algorithm

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(text2vec, dplyr, qlcMatrix, kernlab, knitr)

```

## Load Data

```{r}
load("../output/MS_test.RData")
load("../output/MS_train.RData")

test_data <- MS_test
train_data <- MS_train
```

## Train function

```{r}
train_func <- function(data, C = 3, tau = 0.05, ITER = 200){
  # for C >= 4, phi will become too small that nearly converge to zero
  # thus we choose C = 3 as default
  #data <- train_data
  #C <- 15
  #user <- data[,1]
  #data <- data[,-1]
  
  items <- colnames(data)
  N_item <- nrow(data)
  N_user <- ncol(data)
  #========================
  # Step 1 - initialization
  #========================
  iter <- 1 #iteration
  check_conv <- Inf
  
  mu <- rep(1/C, C)
  
  gamma <- array(0, dim = c(2, N_user, C))
  gamma_1 = matrix(runif(N_user*C, 0, 1), N_user, C)
  gamma_2 = 1-gamma_1
  gamma[1,,] = gamma_1
  gamma[2,,] = gamma_2
  
  phi <- matrix(NA, nrow = N_item, ncol = C)
  pi <- matrix(NA, nrow = N_item, ncol = C)
  pi_pre <- matrix(0, nrow = N_item, ncol = C)

  #========================  
  # Step 2 - Expectation
  #========================
  while(check_conv > tau & iter < ITER){
    for(i in 1:N_item){ #get phi
      sub_mu_phi <- array(0,C) 
      for(j in 1:N_user) {sub_mu_phi <- sub_mu_phi + log(gamma[data[i,j]+1, j,])}
      phi[i, ] <- exp(sub_mu_phi)
    }
    tmp_1 <- mu * phi
    pi <- tmp_1/apply(tmp_1, 1, sum)
    
    #========================
    # Step 3 - Maximization
    #========================
    mu <- apply(pi, 2, sum)/N_item

    for(c in 1:C){ #for gamma, when k = 1
      for(j in 1:N_user){
        sub_sum_pi <- pi[, c] %*% data[, j]
        gamma[2, j, c] <- sub_sum_pi/sum(pi[ ,c])
        gamma[1, j, c] <- 1 - gamma[2, j, c] 
      }
    }
    
    check_conv <- norm(pi - pi_pre)
    print(check_conv)
    print(iter)
    
    pi_pre <- pi
    iter = iter + 1
  }
    
  return(list("pi" = pi, "gamma" = gamma, "mu" = mu, "items" = items))
}  
```

## Prediction Function

```{r}
#Generating prediction
pred_func <- function(data, gamma, mu, pi, items){
  data[is.na(data)] <- 0
  item_test <- colnames(data)
  order <- match(item_test, items)
  
  N_item <- nrow(data)
  N_user <- ncol(data)
  predictions <- as.data.frame(matrix(NA, nrow = N_item, ncol = N_user))
  names(predictions) <- item_test

  user.cluster <- apply(pi, 1, which.max)
  error <- 0
  
  for(i in 1:N_item){
    for(j in 1:ncol(predictions)){
      cluster <- user.cluster[i]
      index <- order[j]
      predictions[i,j] <- which.max(gamma[, index, cluster]) - 1
    }
  }
  
  save(predictions, file = "../data/cluster_prediction.Rdata")
  return(predictions)
}
```

## Ranking score

```{r}
rank_score <- function(predicted_test,true_test){
  ## function to calculate rank score of predicted value
  ## input: predicted_test - predicted value matrix of test data
  ##        true_test - test data matrix
  ## output: rank score
  d <- 0.02
  rank_mat_pred <- ncol(predicted_test)+1-t(apply(predicted_test,1,function(x){return(rank(x,ties.method = 'first'))}))
  rank_mat_test <- ncol(true_test)+1-t(apply(true_test,1,function(x){return(rank(x,ties.method = 'first'))}))
  vec = ifelse(true_test - d > 0, true_test - d, 0)
  R_a <- apply(1/(2^((rank_mat_pred-1)/4)) * vec,1,sum)
  R_a_max <- apply(1/(2^((rank_mat_test-1)/4)) * vec,1,sum)
  
  R <- 100*sum(R_a)/sum(R_a_max)
  return(R)
}
```


```{r}
cluster_train <- train_func(train_data, C = 3)
  
gamma <- cluster_train$gamma
mu <- cluster_train$mu
pi <- cluster_train$pi
items <- cluster_train$items

cluster_pred <- pred_func(test_data, gamma, mu, pi, items)
rank_score(predicted_test = cluster_pred,true_test = test_data)
```






